Approximately Exact Calculations for Linear Mixed Models

### 2015-09-21 ms.tex
splitbox

Let $p$ be an arbitrary point inside $B_1$, as illustrated in Figure~(\ref{fig:boundingbox2}), and define $L \equiv \logRL(p)$; we use $L$ as a lower bound on $\logRL(\sshat)$.  Let $\widetilde{q_1} = (0,\widetilde{\sigma_s^2})$, the intercept of $B$ and the $\sigssq$ axis; let $\widetilde{q_2} = (\sigma_e^{2*},\widetilde{\sigma_s^2})$; let $\widetilde{q_3} = (\widetilde{\sigma_e^2}, \sigma_s^{2*})$; let $\widetilde{q_4} = (\widetilde{\sigma_e^2},0)$, the intercept of $B$ and the $\sigssq$ axis; and let $\logRL_j$ denote the $j$'th term of \eqref{eq:sum}.  The proof considers in turn four regions $Q_1$, $Q_2$, $Q_3$, and $Q_4$ whose union is $B^c$.  Implicitly, $B$ and $B^c$ are functions of $\widetilde{\sigma_e^2}$ and $\widetilde{\sigma_s^2}$.  The proof shows that $\widetilde{\sigma_e^2}$ and $\widetilde{\sigma_s^2}$ can be chosen large enough so that \textbf{D1} is satisfied on each region.  

@@

\item A \textcompute{box} $b$ is a list consisting of
	 \begin{itemize}
	 \item upper and lower limits on $\sigesq$,
	 \item upper and lower limits on $\sigssq$,
	 \item upper and lower bounds on $\logRL$, $U^b$ and $L^b$ from \eqref{eq:bounds},
	 \item indicators for whether this box lies above (or to the right of), below (or to the left of),
	   or straddles each of the $s_z+2$ lines.
	 \end{itemize}
\item \textcompute{killfunc} is a function, shown below \textcompute{findf}, that determines whether
	a box $b$ should be divided more finely.
\item \textcompute{splitbox} is a function that takes a box $b =
	[\sigesq{}_\text{low}, \sigesq{}_\text{high}] \times [\sigssq{}_\text{low}, \sigssq{}_\text{high}]$ as input 
	and returns the four boxes created by dividing each side of $b$ at its midpoint.
\end{enumerate}
\begin{verbatim}
findf <- function(lines, startbox, eps = 0, delE = 0, delS = 0, M = Inf, maxit = 10,
                  ratio = FALSE, lognote = "summary") {

@@

    tmp <- t(vapply(X = c(active, inactive), FUN = function(box) {
        c(box$lims.sigsqs, box$lims.sigsqe, box$bounds)
    }, FUN.VALUE = c(sigsqs.lo = 0.1, sigsqs.hi = 0.1, sigsqe.lo = 0.1, sigsqe.hi = 0.1,
        rll.lower = 0.1, rll.upper = 0.1)))
    end_time <- Sys.time()
    write(paste(lognote, "runtime: ", round(end_time - start_time, digits = 3)),
          file = "log.out", ncolumns = 1, append = TRUE)
    return(data.frame(tmp))
}

# conditions under which a box becomes inactive
killfunc <- function(box, lb, M, eps, delE, delS, ratio) {
    # lb is a global (within startbox) lower bound on max(log(f));
    #      it changes at each iteration.
    #  M, eps, delE, delS stay constant throughout the iterations.
    cond.low <- box$bounds[2] < lb - M
    cond.eps <- diff(box$bounds) < eps
    cond.E <- ifelse(ratio, diff(log(box$lims.sigsqe)) < delE,
                            diff(box$lims.sigsqe) < delE)
    cond.S <- ifelse(ratio, diff(log(box$lims.sigsqs)) < delS,
                            diff(box$lims.sigsqs) < delS)
    return(cond.low || cond.eps || cond.E || cond.S)  # stop dividing this box?
}
\end{verbatim}

>>>>

Let $p$ be an arbitrary point inside $B_1$, as illustrated in Figure~(\ref{fig:boundingbox2}), and define $L \equiv \logRL(p)$; we use $L$ as a lower bound on $\logRL(\sshat)$.  Let $\widetilde{q_1} = (0,\widetilde{\sigma_s^2})$, the intercept of $B$ and the $\sigssq$ axis; let $\widetilde{q_2} = (\sigma_e^{2*},\widetilde{\sigma_s^2})$; let $\widetilde{q_3} = (\widetilde{\sigma_e^2}, \sigma_s^{2*})$; let $\widetilde{q_4} = (\widetilde{\sigma_e^2},0)$, the intercept of $B$ and the $\sigesq$ axis; and let $\logRL_j$ denote the $j$'th term of \eqref{eq:sum}.  The proof considers in turn four regions $Q_1$, $Q_2$, $Q_3$, and $Q_4$ whose union is $B^c$.  Implicitly, $B$ and $B^c$ are functions of $\widetilde{\sigma_e^2}$ and $\widetilde{\sigma_s^2}$.  The proof shows that $\widetilde{\sigma_e^2}$ and $\widetilde{\sigma_s^2}$ can be chosen large enough so that \textbf{D1} is satisfied on each region.  

@@

\item A \textcompute{box} $b$ is a list consisting of
	 \begin{itemize}
	 \item upper and lower limits on $\sigesq$,
	 \item upper and lower limits on $\sigssq$,
	 \item upper and lower bounds on $\logRL$, $U^b$ and $L^b$ from \eqref{eq:bounds},
	 \item indicators for whether this box lies above (or to the right of), below (or to the left of),
	   or straddles each of the $s_z+2$ lines.
	 \end{itemize}
\item \textcompute{killfunc} is a function, shown below \textcompute{findf}, that determines whether
	a box $b$ should be divided more finely.
\item \textcompute{splitbox} is a function that takes a box $b =
	[\sigesq{}_\text{low}, \sigesq{}_\text{high}] \times [\sigssq{}_\text{low}, \sigssq{}_\text{high}]$ as input 
	and returns the four boxes created by dividing each side of $b$ at its midpoint.
\end{enumerate}

\paragraph{\textcompute{findf}: The main function}
\begin{verbatim}
findf <- function(lines, startbox, eps = 0, delE = 0, delS = 0, M = Inf, maxit = 10,
                  ratio = FALSE, lognote = "summary") {

@@

    tmp <- t(vapply(X = c(active, inactive), FUN = function(box) {
        c(box$lims.sigsqs, box$lims.sigsqe, box$bounds)
    }, FUN.VALUE = c(sigsqs.lo = 0.1, sigsqs.hi = 0.1, sigsqe.lo = 0.1, sigsqe.hi = 0.1,
        rll.lower = 0.1, rll.upper = 0.1)))
    end_time <- Sys.time()
    write(paste(lognote, "runtime: ", round(end_time - start_time, digits = 3)),
          file = "log.out", ncolumns = 1, append = TRUE)
    return(data.frame(tmp))
}
\end{verbatim}

\paragraph{\textcompute{killfunc}: determine whether to make a box inactive}
\begin{verbatim}
# conditions under which a box becomes inactive
killfunc <- function(box, lb, M, eps, delE, delS, ratio) {
    # lb is a global (within startbox) lower bound on max(log(f));
    #      it changes at each iteration.
    #  M, eps, delE, delS stay constant throughout the iterations.
    cond.low <- box$bounds[2] < lb - M
    cond.eps <- diff(box$bounds) < eps
    cond.E <- ifelse(ratio, diff(log(box$lims.sigsqe)) < delE,
                            diff(box$lims.sigsqe) < delE)
    cond.S <- ifelse(ratio, diff(log(box$lims.sigsqs)) < delS,
                            diff(box$lims.sigsqs) < delS)
    return(cond.low || cond.eps || cond.E || cond.S)  # stop dividing this box?
}
\end{verbatim}

\paragraph{\textcompute{splitbox}: split a box into four children}
\begin{verbatim}
splitbox <- function(box, lines) {
    # split a box into four children
    NW <- with(box, makebox(lims.sigsqe = c(lims.sigsqe[1], mean(lims.sigsqe)),
        lims.sigsqs = c(mean(lims.sigsqs), lims.sigsqs[2]), status = status,
        lines = lines))

    NE <- with(box, makebox(lims.sigsqe = c(mean(lims.sigsqe), lims.sigsqe[2]),
        lims.sigsqs = c(mean(lims.sigsqs), lims.sigsqs[2]), status = status,
        lines = lines))

    SW <- with(box, makebox(lims.sigsqe = c(lims.sigsqe[1], mean(lims.sigsqe)),
        lims.sigsqs = c(lims.sigsqs[1], mean(lims.sigsqs)), status = status,
        lines = lines))

    SE <- with(box, makebox(lims.sigsqe = c(mean(lims.sigsqe), lims.sigsqe[2]),
        lims.sigsqs = c(lims.sigsqs[1], mean(lims.sigsqs)), status = status,
        lines = lines))

    return(list(NW, NE, SW, SE))
}

\end{verbatim}
